\Large\bf FlexibleSUSY --- A spectrum generator generator for supersymmetric models

### 2014-11-04 doc/flexiblesusy-paper.tex
describe used SG settings for the speed-test
The CKM matrix is set to unity, CP-violating phases are set to zero,
R-parity violation is disabled.

To check the correctness of \fs's generated spectrum generators
extensive unit testing against Softsusy's MSSM and NMSSM
implementations (both $Z_3$-invariant and $Z_3$-violating variants)
was carried out.  These unit tests systematically compare all
tree-level mass matrices, EWSB equations, one- and two-loop
$\beta$-functions, one- and two-loop self-energies and one- and
two-loop tadpoles numerically for certain MSSM and NMSSM parameter
points.  All tested expressions were found to agree within double
machine precision.\footnote{Due to the systematic and detailed tests
  several bugs in Softsusy, SARAH and \fs could be identified and
  corrected.}  Furthermore, the overall pole mass spectrum after the
full fixed-point iteration has finished is compared, and was found to
agree at the sub-permille level.  The origin of the sub-permille level
difference between \fs and Softsusy is the different determination of
the \DRbar\ weak mixing angle
$\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$: \fs calculates
$\theta_W$ from $M_W^2$ and $M_Z^2$ as described in
\secref{sec:calculation-of-gauge-couplings}, while Softsusy determines
$\theta_W$ from the muon decay constant $G_\mu$, which is more
precise.  Furthermore, in Softsusy $3.5.0$ some three-loop
$\beta$-functions and two-loop threshold corrections for the gauge and
Yukawa couplings can be enabled in the MSSM to increase the accuracy
of the RG running and the determination of the \DRbar\ gauge and
Yukawa couplings at $M_Z$ \cite{Allanach:2014nba}.  These corrections
are not implemented in \fs so far.

The complete set of unit tests is shipped with \fs and can be found in
the \code{test/} directory.  The tests can be run with the command
\code{make execute-tests}.  All unit tests are carried out nightly in
order to continuously check the correctness of the meta code and the
generated spectrum generators.  The test results can be found at
\url{http://iktp.tu-dresden.de/~avoigt/FlexibleSUSY/test.xhtml}

@@

We also compared the run-time of \fs against SPheno, Softsusy and the
SARAH generated MSSM spectrum generator SPhenoMSSM.  The test results
can be found in \secref{sec:run-time comparison}.

@@

One of \fs's design goals is a short run-time.  In this section we
demonstrate that this goal was achieved by comparing the run-time of
two different sets of CMSSM spectrum generators:
%
\begin{itemize}
\item \emph{Without sfermion flavour violation:} Disallowing sfermion
  flavour violation simplifies the calculation of the pole masses,
  because flavour-off-diagonal sfermion self-energy matrix elements
  don't need to be calculated.  Here we compare \fs's non-flavour
  violating CMSSM spectrum generator FlexibleSUSY-NoFV (version 1.0.0)
  against SPheno (version 3.2.4) and Softsusy (version 3.4.0).
%
\item \emph{With sfermion flavour violation:} Allowing for sfermion
  flavour violation in general increases the run-time of spectrum
  generators, because the full $6\times 6$ sfermion self-energy
  matrices have to be calculated.  Here we compare FlexibleSUSY-FV
  (version 1.0.0) and SPhenoMSSM (generated with SARAH 4.1.0 and
  linked against SPheno 3.2.4).  Both spectrum generators are based on
  SARAH's MSSM model file, which allows for sfermion flavour
  violation.
\end{itemize}
%
\fs and Softsusy are compiled with g++ 4.8.0 and Intel ifort
13.1.3 20130607.  SPheno and SPhenoMSSM are compiled with Intel ifort
13.1.3 20130607.\footnote{Intel's ifort compiler decreases the
  run-time of SPheno and SPhenoMSSM by approximately a factor $1.5$,
  compared to gfortran.}

For the run-time comparison we're generating $2\cdot 10^{4}$ random
CMSSM parameter points with $m_0\in [50,1000]\unit{GeV}$, $m_{1/2}\in
[50,1000]\unit{GeV}$, $\tan\beta\in [1,100]$, $\sign\mu\in \{-1,+1\}$
and $A_0\in [-1000,$ $1000]\unit{GeV}$.  For each point an SLHA input
file is created by appending the values of $m_0$, $m_{1/2}$,
$\tan\beta$, $\sign\mu$, $A_0$ in form of a \code{MINPAR} block to the
SLHA template file given in
\ref{sec:speed-test-slha-template-file}.  The resulting SLHA input
file is passed to each spectrum generator and the (wall-clock) time is
measured until the program has finished.  The average run-times for
three different CPU types can be found in
\tabref{tab:run-time-comparison}.  The first column shows the run-time
on a Intel Core2 Duo (P8600, $2.40\unit{GHz}$) where only one core was
enabled.  The second column shows the run-time on the same processor
where both cores were enabled.  In the third column a machine with two
Intel Xeon CPUs (L5640, $2.27\unit{GHz}$, $6$ cores) was used.
%
\begin{table}[tbh]
  \centering
  \begin{tabular}{llll}
    \toprule
                            & Intel Core2 Duo    & Intel Core2 Duo   & $2$ $\times$ Intel Xeon\\
                            & (P8600, $1$ core)  & (P8600, $2$ cores)& (L5640, $6$ cores)\\
    \midrule
    FlexibleSUSY-NoFV 1.0.0 & $0.086\unit{s}$    & $0.079\unit{s}$   & $0.060\unit{s}$\\
    SPheno 3.2.4            & $0.119\unit{s}$    & $0.114\unit{s}$   & $0.101\unit{s}$\\
    Softsusy 3.4.0          & $0.175\unit{s}$    & $0.171\unit{s}$   & $0.147\unit{s}$\\
    \midrule
    FlexibleSUSY-FV 1.0.0   & $0.150\unit{s}$    & $0.113\unit{s}$   & $0.074\unit{s}$\\
    SPhenoMSSM 4.1.0        & $0.415\unit{s}$    & $0.401\unit{s}$   & $0.370\unit{s}$\\
    \bottomrule
  \end{tabular}
  \caption{Average run-time of CMSSM spectrum generators
    for random parameter points.  The first three rows show
    spectrum generators which disallow sfermion flavour violation.  Rows
    $4$--$5$ contain sfermion flavour violating spectrum generators, based
    on SARAH's MSSM model file.}
  \label{tab:run-time-comparison}
\end{table}

>>>>

To check the correctness of \fs's generated spectrum generators
extensive unit testing against Softsusy's MSSM and NMSSM
implementations (both $Z_3$-invariant and $Z_3$-violating variants)
has been carried out.  These unit tests systematically compare all
tree-level mass matrices, EWSB equations, one- and two-loop
$\beta$-functions, one- and two-loop self-energies and one- and
two-loop tadpoles numerically for various MSSM and NMSSM parameter
points.  All tested expressions were found to agree within double
machine precision.\footnote{Due to the systematic and detailed tests
  several bugs in Softsusy, SARAH and \fs could be identified and
  corrected.}  Furthermore, the overall pole mass spectrum after the
full fixed-point iteration has finished has been compared, and was
found to agree at the sub-permille level.  The origin of the
sub-permille level difference between \fs and Softsusy is the
different determination of the weak mixing angle
$\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$ in the SUSY model in the
\DRbar\ scheme: \fs calculates
$\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$ from $M_W$ and $M_Z$, as
described in \secref{sec:calculation-of-gauge-couplings}, while
Softsusy determines $\theta_{W,\text{susy}}^{\text{\DRbar}}(M_Z)$ from
the muon decay constant $G_\mu$.  The approach used in Softsusy
results in more precise \DRbar\ gauge couplings at the $M_Z$ scale,
because the muon decay constant is known with a higher accuracy than
the $W$-boson mass.  Furthermore, in Softsusy $3.5.0$ some three-loop
$\beta$-functions and two-loop threshold corrections can be enabled in
the MSSM to increase the accuracy of the RG running and the
determination of the \DRbar\ gauge and Yukawa couplings at $M_Z$
\cite{Allanach:2014nba}.  These corrections are not implemented in \fs
so far.  The complete set of unit tests is shipped with \fs and can be
found in the \code{test/} directory.  The tests can be run with the
command \code{make execute-tests}.  All unit tests are carried out
nightly in order to continuously check the correctness of the meta
code and the generated spectrum generators for the shipped models.
The nightly test results can be found at
\url{http://iktp.tu-dresden.de/~avoigt/FlexibleSUSY/test.xhtml}

@@

We also compared the run-time of \fs against SPheno, Softsusy and the
SARAH generated MSSM spectrum generator SPhenoMSSM.  The results of
the comparison can be found in \secref{sec:run-time comparison}.

@@

One of \fs's design goals is a short run-time.  In this section we
demonstrate that this goal was achieved by comparing the run-time of
two different sets of CMSSM spectrum generators:
%
\begin{itemize}
\item \emph{Without sfermion flavour violation:} Disallowing sfermion
  flavour violation simplifies the calculation of the pole masses,
  because flavour-off-diagonal sfermion self-energy matrix elements
  don't need to be calculated.  Here we compare \fs's non-flavour
  violating CMSSM spectrum generator FlexibleSUSY-NoFV (version 1.0.0)
  against SPheno (version 3.2.4) and Softsusy (version 3.4.0).
%
\item \emph{With sfermion flavour violation:} Allowing for sfermion
  flavour violation in general increases the run-time of spectrum
  generators, because the full $6\times 6$ sfermion self-energy
  matrices have to be calculated.  Here we compare FlexibleSUSY-FV
  (version 1.0.0) and SPhenoMSSM (generated with SARAH 4.1.0 and
  linked against SPheno 3.2.4).  Both spectrum generators are based on
  SARAH's MSSM model file, which allows for sfermion flavour
  violation.
\end{itemize}
%
For the run-time comparison the CKM matrix is set to unity, all CP
violating phases are set to zero and R-parity violation is disabled.
\fs and Softsusy are compiled with g++ 4.8.0 and Intel ifort 13.1.3
20130607.  SPheno and SPhenoMSSM are compiled with Intel ifort 13.1.3
20130607.\footnote{Intel's ifort compiler decreases the run-time of
  SPheno and SPhenoMSSM by approximately a factor $1.5$, compared to
  gfortran.}  We're generating $2\cdot 10^{4}$ random CMSSM parameter
points with $m_0\in [50,1000]\unit{GeV}$, $m_{1/2}\in
[50,1000]\unit{GeV}$, $\tan\beta\in [1,100]$, $\sign\mu\in \{-1,+1\}$
and $A_0\in [-1000,$ $1000]\unit{GeV}$.  For each point an SLHA input
file is created by appending the values of $m_0$, $m_{1/2}$,
$\tan\beta$, $\sign\mu$, $A_0$ in form of a \code{MINPAR} block to the
SLHA template file given in \ref{sec:speed-test-slha-template-file}.
The resulting SLHA input file is passed to each spectrum generator and
the (wall-clock) time is measured until the program has finished.  The
average run-times for three different CPU types can be found in
\tabref{tab:run-time-comparison}.  The first column shows the run-time
on a Intel Core2 Duo (P8600, $2.40\unit{GHz}$) where only one core was
enabled.  The second column shows the run-time on the same processor
where both cores were enabled.  In the third column a machine with two
Intel Xeon CPUs (L5640, $2.27\unit{GHz}$, $6$ cores) was used.
%
\begin{table}[tbh]
  \centering
  \begin{tabular}{llll}
    \toprule
                            & Intel Core2 Duo    & Intel Core2 Duo   & $2$ $\times$ Intel Xeon\\
                            & (P8600, $1$ core)  & (P8600, $2$ cores)& (L5640, $6$ cores)\\
    \midrule
    FlexibleSUSY-NoFV 1.0.0 & $0.086\unit{s}$    & $0.079\unit{s}$   & $0.060\unit{s}$\\
    SPheno 3.2.4            & $0.119\unit{s}$    & $0.114\unit{s}$   & $0.101\unit{s}$\\
    Softsusy 3.4.0          & $0.175\unit{s}$    & $0.171\unit{s}$   & $0.147\unit{s}$\\
    \midrule
    FlexibleSUSY-FV 1.0.0   & $0.150\unit{s}$    & $0.113\unit{s}$   & $0.074\unit{s}$\\
    SPhenoMSSM 4.1.0        & $0.415\unit{s}$    & $0.401\unit{s}$   & $0.370\unit{s}$\\
    \bottomrule
  \end{tabular}
  \caption{Average run-time of CMSSM spectrum generators
    for random parameter points.  The first three rows show
    spectrum generators which disallow sfermion flavour violation.  Rows
    $4$--$5$ contain sfermion flavour violating spectrum generators, based
    on SARAH's MSSM model file.}
  \label{tab:run-time-comparison}
\end{table}
